Range Analysis of Whole Programs

### 2012-03-13 doc/Manuscript/paper.tex
Added some text to the section about experiments.

@@

@@

In this presentation we chose to omit correctness proofs for our algorithms.
For a proof that the widening and the narrowing operators give origin to
approximating sequences, we recommend Cousot and Cousot's work~\cite{Cousot}.
For a proof that the e-SSA form transformation is semantics preserving, we
invite the reader to check a recent report of ours~\cite{XXXX}.
In this work we also show that the results that we obtain via the
sparse analysis are equivalent to the results provided by a dense analysis.
In other words, if the dense analysis tells us that variable $v$ is
associated with the interval range $[l, u]$ at program point $i$, and $v'$
is the new name of $v$, alive at $i$ in the e-SSA form program, then $v'$ is
associated with the interval $[l, u]$ along its entire live range.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We have implemented our range analysis algorithm in LLVM 3.0, and have run
experiments on a Intel quad core CPU with a 2.40GHz clock, and 3.6GB of RAM.
Each core has a 4,096KB L1 cache.
The operation system is Linux Ubuntu 10.04.4.
Our implementation of range analysis has 3,958 lines of
commented C++ code, and our e-SSA conversion module has 672 lines.
We have tested it on 428 C programs that constitute the LLVM test suite, plus
the C programs from SPEC CPU 2006.
These programs together gave us approximately 4.76 million assembly
instructions to analyze.

\noindent
\textbf{Structural Properties.}
Table~\ref{tab:struct} shows structural properties of the programs in
the SPEC 2006 benchmark suite.
Most of the strongly connected components that constitute
the constraint graphs are singletons.
For instance, 99.11\% of the SSCs found in gcc (\texttt{403.gcc}) have only
one node.
Therefore the most complex phases of our algorithm are
confined to a few parts of the constraint graphs that we have to handle.
Furthermore, the strongly connected components are not large -- our largest
component, found in \texttt{403.gcc} has 2,131 nodes, and exists due to a long
chain of mutually recursive function calls.
These two observations imply that our runtime is likely to be linear in
practice, as we show in Figure~\ref{fig:TimeCorr}.

\begin{table}[htdp]
\begin{center}
\begin{scriptsize}
\renewcommand{\tabcolsep}{0.1cm}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|} \hline
Name	    & 464 &	473 & 483 & 458 & 429 & 471 & 403 & 445 & 462 & 401 & 456 \\ \hline
\#I  & 139626 & 8900 & 593765 & 26947 & 2738 & 94036 & 887191 & 141330 &  6237 & 17663 & 61736 \\ \hline
\#$N_v$ & 97494 & 5490 & 352423 & 14407 & 1853	 & 50095 & 449442 & 84846 & 3293 & 12517 & 38409 \\ \hline
\#S  & 91731 & 5108 & 346451 & 13309 & 1751 & 48784 & 435372 & 79909 & 3003 & 11539 & 35619 \\ \hline
L    & 38 & 49 & 120 & 116 & 16 & 258 & 2131 & 311 & 9 & 18 & 41  \\ \hline
\#U  & 89461 & 4983 & 344132 & 12919 & 1726 & 48400 & 431703 & 78407 & 2885 & 11267 & 34631 \\ \hline
\end{tabular}
\end{scriptsize}
\end{center}
\caption{\label{tab:struct}Structural Properties of the programs in the SPEC CPU
2006 benchmark suite.
\#I: number of assembly instructions in the original program.
\#$N_v$: number of variable nodes in the constraint graph.
% TODO: add also number of constraint nodes.
\#S: number of strongly connected components (SCC).
L: size of the largest SCC.
\#U: number of SCCs with only one node.}
\end{table}%

@@

\noindent
\textbf{Runtime Behavior.}
Figure~\ref{fig:TimeCorr} provides a visual comparison between the time to
run our algorithm and the size of the input programs.
We show data for the 100 largest benchmarks in our test suite, in number
of variable nodes in the constraint graph.
We perform function inlining before running our analysis, to increase program
sizes.
Each point in the X line corresponds to a benchmark.
We analyze the smallest benchmark in this set, \texttt{Prolangs-C}, which has
1,131 variable nodes in the constraint graph, in 20ms.
We take 15,91secs to analyze our largest benchmark, \texttt{403.gcc}, which,
after function inlining, has 1,266,273 assembly instructions, and a
constraint graph with 679,652 variable nodes.
For this data set, the coefficient of determination $(R^2)$ is 0.967, which
provides very strong evidence about the linear asymptotic complexity of our
implementation.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=1\textwidth]{images/TimeCorr}
%\includegraphics[width=0.4\textwidth]{images/Complexity}
\end{center}
\caption{\label{fig:TimeCorr}
Correlation between program size (measured in number of nodes in constraint
graphs after inlining) and analysis runtime (ms).
Coefficient of determination = 0.967.
}
\end{figure}

@@

In this presentation we chose to omit correctness proofs for our algorithms.
For a proof that the widening and the narrowing operators give origin to
approximating sequences, we recommend Cousot and Cousot's work~\cite{Cousot77}.
For a proof that the e-SSA form transformation is semantics preserving, we
invite the reader to check a recent report of ours~\cite{XXXX}.
In this work we also show that the results that we obtain via the
sparse analysis are equivalent to the results provided by a dense analysis.
In other words, if the dense analysis tells us that variable $v$ is
associated with the interval range $[l, u]$ at program point $i$, and $v'$
is the new name of $v$, alive at $i$ in the e-SSA form program, then $v'$ is
associated with the interval $[l, u]$ along its entire live range.
